SFJ: An implementation of Semantic Featherweight Java

### 2020-02-27 conference/paper.tex
types and terms

\section{SFJ: Design and Implementation}
\label{sec:design}
As already discussed in the introduction, the key idea behind semantic subtyping is interpreting types as sets, in particular,  as sets of values of the language at hand. Since we want these types to be able to be finitely representable, we restrict our types to finite trees whose leaves are constants with no cycles. 
For example a recursive type $\alpha = [a : \alpha]$ would be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)). 
Similarly the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.

Therefore, when processing the abstract syntax tree (AST) of a program in SFJ, we do not allow the type of the class we are defining to be a type of its own field. 
Secondly we mark any classes with only basic types \emph{(bool, int, float ...)} for the types of its fields as \emph{resolved} and otherwise, if it contains a class type, we mark it as \emph{unresolved}. 
After processing the whole AST, we perform the following algorithm to decide whether the type definitions in the program are valid.
\begin{verbatim}
do
    boolean resolutionOccured = false
    for class that is unresolved:
        boolean resolved = true
        for field that contains a class type:
            if class type is not resolved
                resolved = false
        if resolved is true:
            set class to be resolved
            resolutionOccured = true
while resolutionOccured is true

>>>>

\section{SFJ: Design and Implementation}
\label{sec:design}
The syntax of types is given by the following grammar \cite{}:
$$
\begin{array}{llll}
\tau & ::= & \alpha \ |\ \mu
& \mbox{{\em Type term}}
\\
\alpha & ::=  & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha \synwedge \alpha \ |\ \synneg\alpha
\qquad
& \mbox{{\em Object type} ($\alpha $-type)}
\\
\mu & ::=  & \alpha \to \alpha \ |\ \mu \synwedge \mu \ |\ \synneg\mu
& \mbox{{\em Method type} ($\mu $-type)}
\end{array}
$$
$\alpha$-types are used to type fields and $\mu$-types are used to type methods.
Type $\zero$ is the empty type.
Type $\btypes$ denotes the \emph{basic} types, such as integers, booleans, etc.
{Record} types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels are used to type objects.
Arrow types $\alpha \to \alpha$ are used to type methods.
The boolean types using $\synwedge$ and $\synneg$ have their expected meanings and $\synvee$ is obtained by their combination.\\
The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{}:
\begin{align*}
& \mbox{\textit{Class declaration}} & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }\\
& \mbox{\textit{Constructor}} & K \;::=\; & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \}\\
& \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \} \\
& \mbox{\textit{Expressions}} & e \; ::=\; & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
\end{align*}
Since we want to use types $\tau$ in practice in SFJ, we restrict them to finite trees whose leaves are constants with no cycles.
For e.g., a recursive type $\alpha = [a : \alpha]$ would be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we cannot instate it.
%Similarly the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.

Therefore, when processing the abstract syntax tree (AST) of a program in SFJ, we do not allow the type of the class we are defining to be a type of its own field. 
Secondly, we mark any classes with fields only of basic types as \emph{resolved} and otherwise, if it contains a record type, we mark it as \emph{unresolved}. 
After processing the whole AST, we perform the following algorithm to decide whether the type definitions in the program are valid.
\begin{verbatim}
do
    boolean resolutionOccured = false
    for class that is unresolved:
        boolean resolved = true
        for field that contains a class type:
            if class type is not resolved
                resolved = false
        if resolved is true:
            set class to be resolved
            resolutionOccured = true
while resolutionOccured is true
